
```{r}
library(rpart)
library(rpart.plot)
library(caret)
library(pROC)
library(randomForest) #for randm forest
library(ggplot2)
library(dplyr)
library(gridExtra)
```

```{r}
thyroid=read.csv("thyroid_diff.csv")
summary(thyroid)
```

```{r}
str(thyroid)
```

```{r}
thyroid$Gender <- factor(thyroid$Gender)
thyroid$Smoking <- factor(thyroid$Smoking)
thyroid$Hx.Smoking <- factor(thyroid$Hx.Smoking)
thyroid$Hx.Radiothreapy <- factor(thyroid$Hx.Radiothreapy)
thyroid$Thyroid.Function <- factor(thyroid$Thyroid.Function)
thyroid$Physical.Examination <- factor(thyroid$Physical.Examination)
thyroid$Adenopathy <- factor(thyroid$Adenopathy)
thyroid$Pathology <- factor(thyroid$Pathology)
thyroid$Focality <- factor(thyroid$Focality)
thyroid$Risk <- factor(thyroid$Risk)
thyroid$T <- factor(thyroid$T)
thyroid$N <- factor(thyroid$N)
thyroid$M <- factor(thyroid$M)
thyroid$Stage <- factor(thyroid$Stage)
thyroid$Response <- factor(thyroid$Response)
thyroid$Recurred <- factor(thyroid$Recurred)
```

```{r}
str(thyroid)
```
```{r}
sum(is.na(thyroid)) 
```

```{r}
vnames=c("Age","Gender","Smoking","Hx.Smoking","Hx.Radiothreapy", 
         "Thyroid.Function", "Physical.Examination", "Adenopathy", "Pathology", 
         "Focality", "Risk", "T", "N", "M", "Stage", "Response","Recurred")
mydat=thyroid
mydat2=mydat[,vnames]
```
```{r}
id=sample(1:nrow(mydat2), size= floor(nrow(mydat2)*0.80))
traindat=mydat2[id,]
testdat=mydat2[-id,]
library(ggplot2)
ggplot(traindat, aes(x = Recurred)) + geom_bar()
```
## EDA

### REcurrence dist
```{r}
R <- ggplot(traindat, aes(x = Recurred, fill = Recurred)) +
  geom_bar() +
  theme_minimal() +
  ggtitle("Recurrence Distribution") +
  xlab("Recurred (Yes/No)") +
  ylab("Count") +
  scale_fill_manual(values = c("skyblue", "salmon")) +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10),
    legend.position = "none"
  )
R
```
### Age Distribution
```{r}
p1 <- ggplot(traindat, aes(x = Age)) +
  geom_histogram(binwidth = 5, fill = "steelblue", color = "black") +
  theme_minimal() +
  ggtitle("Age Distribution") +
  xlab("Age") +
  ylab("Frequency")
p1
```

### Gender Distribution
```{r}
p2 <- ggplot(traindat, aes(x = Gender, fill = Gender)) +
  geom_bar() +
  theme_minimal() +
  ggtitle("Gender Distribution") +
  xlab("Gender") +
  ylab("Count")
p2
```
## Recurred vs Risk
```{r}
RR <- ggplot(traindat, aes(x = Risk, fill = Recurred)) +
  geom_bar(position = "fill") +
  theme_minimal() +
  ggtitle("Recurrence vs. Risk") +
  xlab("Risk") +
  ylab("Proportion") +
  scale_fill_manual(values = c("skyblue", "salmon"))
RR
```

## Smoking history and risk
```{r}
p3 <- ggplot(traindat, aes(x = Hx.Smoking, fill = Risk)) +
  geom_bar(position = "fill") +
  theme_minimal() +
  ggtitle("Smoking History vs. Risk") +
  xlab("Smoking History") +
  ylab("Proportion") +
  scale_fill_brewer(palette = "Set3")
p3
```

### Thyroid Stages
```{r}
p4 <- ggplot(traindat, aes(x = Stage, fill = Thyroid.Function)) +
  geom_bar(position = "dodge") +
  theme_minimal() +
  ggtitle("Thyroid Function by Stage") +
  xlab("Stage") +
  ylab("Count") +
  scale_fill_brewer(palette = "Pastel1")
p4
```

### Recurrence by Response
```{r}
p5 <- ggplot(traindat, aes(x = Response, fill = Recurred)) +
  geom_bar(position = "fill") +
  theme_minimal() +
  ggtitle("Recurrence by Response") +
  xlab("Response") +
  ylab("Proportion") +
  scale_fill_manual(values = c("skyblue", "salmon"))
p5
```
Terlihat orang yang memiliki excellent response, penyakit tiroid lebih banyak tidak kambuh

### Recurred vs N and Adenopathy
```{r}
RNA <- ggplot(traindat, aes(x = N, fill = Recurred)) +
  geom_bar(position = "fill") +
  facet_wrap(~Adenopathy) +
  theme_minimal() +
  ggtitle("Recurrence by N Stage and Adenopathy") +
  xlab("N Stage") +
  ylab("Proportion") +
  scale_fill_manual(values = c("skyblue", "salmon"))
RNA
```

### Recurred vs age and risk
```{r}
RAR <- ggplot(traindat, aes(x = Age, y = Risk, color = Recurred)) +
  geom_jitter(alpha = 0.7) +
  theme_minimal() +
  ggtitle("Recurrence by Age and Risk") +
  xlab("Age") +
  ylab("Risk") +
  scale_color_manual(values = c("darkgreen", "red"))
RAR
```

### Recurred vs Response, Risk, and N
```{r}
RRRN <- ggplot(traindat, aes(x = Response, fill = Recurred)) +
  geom_bar(position = "dodge") +
  facet_grid(Risk ~ N) +
  theme_minimal() +
  ggtitle("Recurrence by Response, Risk, and N") +
  xlab("Response") +
  ylab("Count") +
  scale_fill_manual(values = c("skyblue", "salmon"))
RRRN
```

#### Combine all plot
```{r}
grid.arrange(R,p1, p2, p3, p4, p5, RNA, RAR, RRRN, nrow = 3, ncol=3)
```
chi square table
```{r}
# Load necessary libraries
library(readr)
library(ggplot2)
library(tidyr)

# Load the dataset
thyroid_diff <- read_csv("thyroid_diff.csv")

# Define a function to perform chi-square test and plot histogram
plot_chisquare_histogram <- function(data, variable1, variable2) {
  # Create a contingency table
  contingency_table <- table(data[[variable1]], data[[variable2]])
  
  # Perform chi-square test
  chisq_test <- chisq.test(contingency_table)
  
  # Extract observed and expected frequencies
  observed <- as.vector(chisq_test$observed)
  expected <- as.vector(chisq_test$expected)
  
  # Create a dataframe for plotting
  histogram_data <- data.frame(
    Category = rep(1:length(observed), 2),  # Each category (1 to n) repeated twice
    Frequency = c(observed, expected),
    Type = rep(c("Observed", "Expected"), each = length(observed))
  )
  
  # Create the histogram
  p <- ggplot(histogram_data, aes(x = factor(Category), y = Frequency, fill = Type)) +
    geom_bar(stat = "identity", position = "dodge", color = "black") +
    scale_fill_manual(values = c("skyblue", "orange")) +
    labs(
      title = paste("Observed vs Expected Frequencies for", variable1, "and", variable2),
      x = "Category",
      y = "Frequency",
      fill = "Type"
    ) +
    theme_minimal()
  
  return(p)
}

# Iterate through all possible pairs of variables
for (var1 in colnames(thyroid_diff)) {
  for (var2 in colnames(thyroid_diff)) {
    if (var1 != var2) {
      print(paste("Generating plot for:", var1, "and", var2))
      p <- plot_chisquare_histogram(thyroid_diff, var1, var2)
      print(p)
    }
  }
}

```
-------------------------------------------------------------------------
### Regression Tree
```{r}
summary(thyroid)
```
```{r}
vnames=c("Age","Gender","Smoking","Hx.Smoking","Hx.Radiothreapy", 
         "Thyroid.Function", "Physical.Examination", "Adenopathy", "Pathology", 
         "Focality", "Risk", "T", "N", "M", "Stage", "Response","Recurred")

```

```{r}
mydat=thyroid
library(ggplot2)
ggplot(traindat, aes(x = Recurred)) + geom_bar()
```
mari membuat model decision tree
```{r}
# Perform grid search for minsplit and maxdepth
library(caret)

# Define the grid of parameters to tune
grid <- expand.grid(minsplit = c(10, 20, 30), maxdepth = c(5, 10, 15))

# Function to train and evaluate models with different parameters
tune_tree <- function(minsplit, maxdepth) {
  tree <- rpart(Recurred ~ ., data = traindat, method = "class",
                control = rpart.control(minsplit = minsplit, maxdepth = maxdepth, cp = 0.01))
  # Use cross-validation to evaluate
  cv_result <- printcp(tree)
  return(tree$cptable[which.min(tree$cptable[, "xerror"]), "xerror"])
}

# Apply grid search
results <- apply(grid, 1, function(params) tune_tree(params[1], params[2]))
best_params <- grid[which.min(results), ]

# Identify the best parameters
best_params <- grid[which.min(results), ]
cat("Best Parameters: minsplit =", best_params$minsplit, ", maxdepth =", best_params$maxdepth, "\n")
library(rpart)
best_minsplit <- 10    # Replace with your optimal minsplit value
best_maxdepth <- 5     # Replace with your optimal maxdepth value

# Build the decision tree with tuned parameters
final_tree <- rpart(Recurred ~ ., 
                    data = traindat, 
                    method = "class", 
                    control = rpart.control(minsplit = best_minsplit, maxdepth = best_maxdepth, cp = 0.01))

# View the tree structure
print(final_tree)
par(mar = c(1, 1, 1, 1)) # Adjust margins
plot(final_tree, uniform = TRUE, margin = 0.05)
text(final_tree, use.n = TRUE, cex = 0.8)

```

```{r}
set.seed(123)
predictions <- predict(final_tree, testdat, type = "class")
confusion_matrix <- table(testdat$Recurred, predictions)
accuracy <- sum(diag(confusion_matrix)) / sum(confusion_matrix)
confusion_matrix
accuracy
```
```{r}
# Assuming 'confusion_matrix' contains your table
confusion_matrix <- table(testdat$Recurred, predictions)

# Extract values from the confusion matrix
TP <- confusion_matrix[2, 2]  # True Positives
FP <- confusion_matrix[1, 2]  # False Positives
FN <- confusion_matrix[2, 1]  # False Negatives
TN <- confusion_matrix[1, 1]  # True Negatives

# Calculate precision, recall, and F1-score
precision <- TP / (TP + FP)  # Precision
recall <- TP / (TP + FN)     # Recall
f1_score <- 2* (precision * recall) / (precision + recall)  # F1-score

# Print results
cat("Precision: ", precision, "\n")
cat("Recall: ", recall, "\n")
cat("F1-score: ", f1_score, "\n")
```
```{r}
# Load necessary libraries
library(ggplot2)
library(reshape2)

# Assuming `confusion_matrix` is already created as a table
# Convert the confusion matrix into a data frame for ggplot
conf_matrix_df <- as.data.frame(as.table(confusion_matrix))
colnames(conf_matrix_df) <- c("Actual", "Predicted", "Frequency")

# Create a heatmap of the confusion matrix
confusion_matrix_plot <- ggplot(conf_matrix_df, aes(x = Predicted, y = Actual, fill = Frequency)) +
  geom_tile(color = "white") +
  geom_text(aes(label = Frequency), color = "black", size = 5) +
  scale_fill_gradient(low = "lightblue", high = "darkblue") +
  labs(
    title = "Confusion Matrix Heatmap",
    x = "Predicted Class",
    y = "Actual Class",
    fill = "Frequency"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(vjust = 0.5)
  )

# Print the confusion matrix plot
print(confusion_matrix_plot)
```

## Random Forest

```{r}
set.seed(123)
forest <- randomForest(Recurred~., data= traindat, proximity= TRUE ) #proximity= TRUE untuk return proximity matrix. digunakan untuk cluster samples
forest
plot(forest)
```

```{r}
oob.error.data <- data.frame(
  Trees= rep(1:nrow(forest$err.rate), times=3),
  Type= rep(c("OOB", "Yes", "No"), each= nrow(forest$err.rate)),
  Error= c(forest$err.rate[,"OOB"], 
           forest$err.rate[,"Yes"],
           forest$err.rate[,"No"])
  ) 

ggplot(data= oob.error.data, aes(x=Trees, y= Error)) + geom_line(aes(color= Type))
```

```{r}
forestcek <- randomForest(Recurred~., data= traindat, ntree=1000, proximity= TRUE )
forestcek
```

```{r}
oob.error.data2 <- data.frame(
  Trees2= rep(1:nrow(forestcek$err.rate), times=3),
  Type2= rep(c("OOB", "Yes", "No"), each= nrow(forestcek$err.rate)),
  Error2= c(forestcek$err.rate[,"OOB"], 
           forestcek$err.rate[,"Yes"],
           forestcek$err.rate[,"No"])
  ) 

ggplot(data= oob.error.data2, aes(x=Trees2, y= Error2)) + geom_line(aes(color= Type2))
```

```{r}
oob.values <- vector(length= 10)
for (i in 1:10) {
  temp.model <- randomForest(Recurred~., data= traindat, mtry=i, ntree=1000)
  oob.values[i] <- temp.model$err.rate[nrow(temp.model$err.rate),1]
}
oob.values
```

```{r}
distance.matrix <- dist(1-forest$proximity)
mds.stuff <- cmdscale(distance.matrix, eig= TRUE, x.ret = TRUE)
mds.var.per <- round(mds.stuff$eig/sum(mds.stuff$eig)*100,1)

mds.values <- mds.stuff$points
mds.data <- data.frame(Sample= rownames(mds.values),
                       X= mds.values[,1],
                       Y= mds.values[,2],
                       Status= traindat$Recurred)

ggplot(data= mds.data, aes(x=X, y=Y, label= Sample))+
  geom_text(aes(color= Status))+
  theme_bw() +
  xlab(paste("MDS1 -", mds.var.per[1], "%", sep=""))+
  ylab(paste("MDS2 -", mds.var.per[2], "%", sep=""))+
  ggtitle("MDS plot using(1 - Random Forest Proximities)")
```

```{r}
predictions2 <- predict(forest, newdata= testdat, type = "class")
confusion_matrix2 <- table(testdat$Recurred, predictions2)
confusion_matrix2
```

```{r}
accuracy2 <- sum(diag(confusion_matrix2)) / sum(confusion_matrix2)
accuracy2
```

```{r}
var.imp <- importance(forest)
var.imp
```

```{r}
order(importance(forest)) #lowest to highest
importance(forest)[rank(importance(forest)),]
```
```{r}
var.imp.df <- data.frame(Variabel= rownames(var.imp), Importance= var.imp[,"MeanDecreaseGini"])
var.imp.df <- var.imp.df[order(var.imp.df$Importance, decreasing = TRUE),]

ggplot(var.imp.df, aes(x = reorder(Variabel, Importance), y = Importance)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  labs(
    title = "Variable Importance Plot",
    x = "Variabel",
    y = "Importance (Mean Decrease Gini)"
  ) +
  theme_minimal()

```


Bayes
```{r}
# Step 1: Install and load e1071 package
library(e1071)

# Step 2: Prepare the data
traindat$Recurred <- as.factor(traindat$Recurred)

# Step 3: Train the Naive Bayes model
nb_model <- naiveBayes(Recurred ~ ., data = traindat)

# Print the Naive Bayes model summary
print(nb_model)

# Step 4: Make predictions on the test set
nb_predictions <- predict(nb_model, newdata = testdat)

# Convert predictions and actual values to factors with the same levels
nb_predictions <- factor(nb_predictions, levels = levels(testdat$Recurred))
testdat$Recurred <- factor(testdat$Recurred)

# Generate the confusion matrix
# Align levels between predictions and actual values
common_levels <- union(levels(nb_predictions), levels(testdat$Recurred))
nb_predictions <- factor(nb_predictions, levels = common_levels)
testdat$Recurred <- factor(testdat$Recurred, levels = common_levels)

# Remove unused levels if necessary
nb_predictions <- droplevels(nb_predictions)
testdat$Recurred <- droplevels(testdat$Recurred)

# Generate confusion matrix
library(caret)
conf_matrix1 <- confusionMatrix(nb_predictions, testdat$Recurred)

# Print the confusion matrix
print(conf_matrix1)


# Step 6: Print model accuracy
cat("Accuracy: ", conf_matrix1$overall["Accuracy"], "\n")

# Optional: Plot the confusion matrix
# Assuming conf_matrix is generated
cm_table <- conf_matrix1$table

# Convert to a data frame
cm_df <- as.data.frame(as.table(cm_table))
colnames(cm_df) <- c("Actual", "Predicted", "Frequency")

# Plot using ggplot2
library(ggplot2)

ggplot(cm_df, aes(x = Predicted, y = Actual, fill = Frequency)) +
  geom_tile(color = "white") +  # Create the heatmap tiles
  geom_text(aes(label = Frequency), color = "black", size = 5) +  # Add text labels
  scale_fill_gradient(low = "white", high = "blue") +  # Define the color gradient
  theme_minimal() +
  labs(
    title = "Confusion Matrix Heatmap",
    x = "Predicted",
    y = "Actual"
  )
print(conf_matrix1)
```
```{r}
library(caret)

# Assuming conf_matrix1 is already created
# Extract precision, recall, and F1-score for the positive class
precision <- conf_matrix1$byClass["Pos Pred Value"]  # Precision
recall <- conf_matrix1$byClass["Sensitivity"]        # Recall
f1_score <- conf_matrix1$byClass["F1"]               # F1-score (if available)

# Print the results
cat("Precision: ", precision, "\n")
cat("Recall: ", recall, "\n")
cat("F1-Score: ", f1_score, "\n")
```
